### 2024-11-18 paper/paper.tex
Add new publication paper/paper.pdf

% A generic article template
\documentclass[11pt]{article}

\usepackage{sectsty}
\usepackage{graphicx}

% Margins
\topmargin=-0.45in
\evensidemargin=0in
\oddsidemargin=0in
\textwidth=6.5in
\textheight=9.0in
\headsep=0.25in

\title{The article}
\author{ The Author }
\date{\today}

\begin{document}
\maketitle
\pagebreak

% Optional TOC
% \tableofcontents
% \pagebreak

%--Paper--

\section{Section 1}

Lorem Ipsum

% \includegraphics{../figures/my-figure.png}

\pagebreak

\section{Section 2}

Lorem Ipsum \\

%--/Paper--

\end{document}